---
title: "DAP_figures"
author: "sophie whikehart"
date: "2023-12-04"
output: 
  word_document:
    keep_md: yes
    toc: TRUE
    toc_depth: 4
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      comment = "")

# load libraries

library(e1071) # skewness()

library(tidyverse) # summarize data and for %>% function

library(ggplot2) # create graphs using the ggplot2 package

library(flextable) # create a nice summary table in Word

library(gtsummary) # summary table in word 

# import data 
chdclass <- read.csv(file = "data/chd3200 (1).csv")

#create factor variables 

# MI family history
chdclass$fhha.f <- factor(chdclass$fhha, levels = c(0,1), 
                        labels = c("No", "Yes"))

# coronary heart disease (CHD)
chdclass$incchd.f <- factor(chdclass$incchd, levels = c(0,1), 
                          labels = c("No", "Yes"))
# smoking status
chdclass$smoke.f <- factor(chdclass$smoke, levels = c(1,2,3), 
                            labels = c("Never", "Former", "Current"))
```

# Figure 1

## Conduct descriptive analysis for the quantitative variables by creating a summary table for the quantitative variables [age, alcoh, hdl, insulin]

```{r, message=FALSE}
chdclass %>% select(age, alcoh, hdl, insulin) %>%
  tbl_summary(missing = "ifany",
              type = all_continuous() ~ "continuous2",
              statistic = all_continuous() ~ c("{mean} ({sd})",
                                                "{median} ({p25}, {p75})", 
                                                "{min}, {max}",
                                                "{skewness}"),
              digits = list(all_continuous() ~ 1),  
              label = list(age ~ "Age (y)",
                           alcoh  ~ "Alcohol consumption (drinks/wk)",
                           hdl ~ "HDL cholesterol (mg/dL)",
                           insulin ~ "Serum insulin (IU/ml)")
              ) %>%
  modify_header(label = "**Variable**") %>% 
  add_stat_label(label = all_continuous() ~ c("Mean (SD)", 
                                              "Median (IQR)",
                                              "Min to Max", "Skewness"))
```

# Figure 2 
# Histogram to show distribution of alcohol

```{r Alcohol histogram, message=FALSE}
ggplot(data=chdclass, aes(x=alcoh)) +
geom_histogram(color="black", fill="blue4", binwidth=10, position="identity", alpha=0.7) +
labs(x="Alcohol drinks",
y="No. of subjects",
title="Histogram of Alcohol Consumption (drinks/wk)") 
``` 

# Figure 3 
# Histogram to show distribution of hdl 

```{r hdl histogram, message=FALSE}
ggplot(data=chdclass, aes(x=hdl)) +
geom_histogram(color="black", fill="lightblue", binwidth=10, position="identity", alpha=0.7) +
labs(x="HDL",
y="No. of subjects",
title="HDL Cholesterol (mg/dL)") 
``` 

```{r, include= TRUE}
#create categories
chdclass$alcoh <- cut(chdclass$alcoh, breaks = c(-Inf,0,7.000,Inf))

#check categories
table(chdclass$alcoh, exclude=FALSE)

#create a factor variable 
chdclass$alcoh.f <- factor(chdclass$alcoh,
                           levels = c("(-Inf,0]", "(0,7]",
                           "(7, Inf]"),
                           labels = c("0", ">0 to 7", ">7"))
#check factor variable
table(chdclass$alcoh.f)
```

# Figure 4
## Histogram to show distribution of serum insulin levels

```{r HDL inaulin, message=FALSE}
ggplot(data=chdclass, aes(x=hdl)) +
geom_histogram(color="black", fill="purple4", binwidth=10, position="identity", alpha=0.7) +
labs(x="Insulin (Serum insulin level (IU/ml))",
y="No. of subjects",
title="Histogram of Insulin") 
```

# Figure 5
## Frequency distribution of categorical variables 

```{r, message=FALSE}
theme_gtsummary_journal(journal = "jama") # Use the JAMA theme

chdclass %>% select(fhha.f, incchd.f, smoke.f, alcoh.f) %>%
  tbl_summary(missing = "ifany",
              type = list(fhha.f ~ "categorical",
                          incchd.f ~ "categorical"),
              digits = list(all_categorical() ~ c(0,1)),  
              label = list(fhha.f ~ "Family history MI",
                           incchd.f ~ "Coronary heart disease",
                           smoke.f ~ "Smoking status",
                           alcoh.f ~ "Average consumption of alcohol per week")
              ) %>%
  modify_header(label = "**Variable**")
```

# Figure 6
## Bar chart of Family History of Heart Attack

```{r FFHA barchart, message=FALSE}
ggplot(data=chdclass, aes(x=fhha.f)) +
geom_bar(color="black", fill="red4", binwidth=10, position="identity", alpha=0.7) +
labs(x="Family History of Heart Attack",
y="No. of subjects",
title="Bar chart on Family History of Heart Attack") 
```

# Figure 7
## Bar chart on smoking status

```{r Smoke barchart, message=FALSE}
ggplot(data=chdclass, aes(x=smoke.f)) +
geom_bar(color="black", fill="black", binwidth=10, position="identity", alpha=0.7) +
labs(x="Smoking Status",
y="No. of subjects",
title="Bar chart on Smoking Status") 
```

# Figure 8
## Bar chart to show alcohol consumption 

```{r alcohol barchart, message=FALSE}
ggplot(data=chdclass, aes(x=alcoh.f)) +
geom_bar(color="black", fill="lightblue4", binwidth=10, position="identity", alpha=0.7) +
labs(x="Alcohol drinks",
y="No. of subjects",
title="Bar chart on Alcohol Consumption by 0 drinks, >0 to 7 and >7 drinks per week") 
```

# Figure 9
## Box plot of CHD and HDL Cholestrol 

```{r, fig.width=3, fig.height=3}
ggplot(data=chdclass, aes(x=incchd.f, y=hdl, fill=incchd.f)) +
  geom_boxplot() +
  labs(x="CHD",
       y="HDL cholesterol (mg/dL)",
       title="HDL cholesterol by CHD") + 
  scale_y_continuous(limits=c(0, 140), breaks=seq(0,150,20)) + 
  theme_classic() +
  theme(legend.position="none") +
   theme(
    legend.position = "none",
    legend.title = element_text(colour="black", size=10, 
                                      face="bold"),
    legend.text = element_text(colour="black", size=10, 
                                     face="bold"),
    axis.title.x = element_text(face="bold", size=12),
    axis.text.x = element_text(face="bold", size=11),
    axis.title.y = element_text(face="bold", size=12),
    axis.text.y = element_text(face="bold", size=11)) +
  scale_fill_manual(values=c("orange","lightblue"))
```

# Figure 10
## Two sample t-test to determine if HDL cholesterol levels are associated with CHD 

```{r }
# t.test 
t.test(hdl ~ incchd.f, conf.level=0.95, var.equal = FALSE, data=chdclass, alternative= "two.sided")
```

# Figure 11
## Final summary table reporting HDl cholesterol, alcohol, serum insulin, smoking status and family history of heart attack by CHD 

```{r, message=FALSE}
theme_gtsummary_journal(journal = "jama") # Use the JAMA theme

chdclass %>% select(incchd.f, age, alcoh.f, hdl, insulin, smoke.f, fhha.f) %>%
  tbl_summary(by = incchd.f,
              missing = "ifany",
              statistic = list(age ~ c("{median} ({p25}, {p75})"),
                               hdl ~ c("{mean} ({sd})"),
                               insulin ~ c("{median} ({p25}, {p75})")
                               ),
              percent = "row",
              type = list(fhha.f ~ "categorical",
                          alcoh.f ~ "categorical"),
              digits = list(all_categorical() ~ c(0,1)),  
              label = list(hdl ~ "HDL cholesterol (mg/dL)",
                           insulin ~ "Serum insulin (IU/ml)",
                           smoke.f ~ "Smoking status",
                           age ~ "Age (years)",
                           alcoh.f ~ "Average number of alcoholic drinks per week",
                           fhha.f ~ "Family history of heart attack")
              ) %>%
  add_overall() %>%
  modify_header(label = "**Variable**") %>%
  modify_spanning_header(stat_1 ~ "Coronary Heart Disease",
                         stat_2 ~ "Coronary Heart Disease") %>% 
  add_p()
```

# Figure 12
## Overall HDL cholesterol by CHD 

```{r}
#categorizing HDL cholesterol
chdclass$hdlcat <- cut(chdclass$hdl, breaks = c(-Inf,40,Inf))

table(chdclass$hdlcat, exclude = FALSE)

# create a factor variable
chdclass$hdlcat.f <- factor(chdclass$hdlcat, 
                            levels = c("(-Inf,40]", "(40, Inf]"), 
                            labels = c("<40", ">=40"))

# check factor variable
table(chdclass$hdlcat.f)

#create cross-tab for overall HDL cholesterol and CHD incidence
chdclass |> 
  select(incchd.f, hdlcat.f) |>
  tbl_summary(by = incchd.f,
              missing = "no",
              percent = "row",
              statistic = all_categorical() ~ "{n} ({p}%)",
              label = list(incchd.f ~ "Incidence of CHD",
                           hdlcat.f ~ "HDL cholesterol")) %>%
  modify_spanning_header(stat_1 ~ "Overall HDL and CHD")

```

# Figure 13
## Bar chart on categorized HDL 

```{r hdl barchart, message=FALSE}
ggplot(data=chdclass, aes(x=hdlcat.f)) +
geom_bar(color="black", fill="pink4", binwidth=10, position="identity", alpha=0.7) +
labs(x="HDL",
y="No. of subjects",
title="Bar chart on HDL <40 and >=40") 
```

# Figure 14
## HDL cholesterol by CHD - Family History of Heart Attack = Yes

```{r}
#categorizing HDL cholesterol
chdclass$hdlcat <- cut(chdclass$hdl, breaks = c(-Inf,40,Inf))

table(chdclass$hdlcat, exclude = FALSE)

#create a factor variable
chdclass$hdlcat.f <- factor(chdclass$hdlcat, 
                            levels = c("(-Inf,40]", "(40, Inf]"), 
                            labels = c("<40", ">=40"))

#check factor variable
table(chdclass$hdlcat.f)

#create cross-tab for FFHA = YES by HDL cholesterol and CHD incidence
chdclass %>% 
  filter(fhha.f == "Yes") %>%
  select(incchd.f, hdlcat.f) %>%
  tbl_summary(by = incchd.f,
              missing = "no",
              percent = "row",
              statistic = all_categorical() ~ "{n} ({p}%)",
              label = list(incchd.f ~ "Incidence of CHD",
                           hdlcat.f ~ "HDL cholesterol (mg/dL)")) %>%
modify_header(label = "**Variable**") %>%
  modify_spanning_header(stat_1 ~ "Family History of Heart Attack = Yes to CHD")
```



# Figure 15
## HDL cholesterol by CHD - Family History of Heart Attack = No

```{r}
#categorizing HDL cholesterol
chdclass$hdlcat <- cut(chdclass$hdl, breaks = c(-Inf,40,Inf))

table(chdclass$hdlcat, exclude = FALSE)

#create a factor variable
chdclass$hdlcat.f <- factor(chdclass$hdlcat, 
                            levels = c("(-Inf,40]", "(40, Inf]"), 
                            labels = c("<40", ">=40"))

#check factor variable
table(chdclass$hdlcat.f)

#create cross-tab for FFHA = NO by HDL cholesterol and CHD incidence
chdclass %>% 
  filter(fhha.f == "No") %>%
  select(incchd.f, hdlcat.f) %>%
  tbl_summary(by = incchd.f,
              missing = "no",
              percent = "row",
              statistic = all_categorical() ~ "{n} ({p}%)",
              label = list(incchd.f ~ "Incidence of CHD",
                           hdlcat.f ~ "HDL cholesterol (mg/dL)")) %>%
modify_header(label = "**Variable**") %>%
  modify_spanning_header(stat_1 ~ "Family History of Heart Attack = No to CHD") 
```

